---
title: "Model tuning"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, eval = FALSE)
```

```{r, echo = FALSE}
library(ampir)
library(tidyverse)
library(caret)
library(pROC)
source("R/remove_nonstandard_aa.R")
```

## Data preparation 

To use the `tune_model.R` command-line script we need to prepare a target and background database that is balanced and free from invalid entries.  By default the minimum length is 20 so we filter both databases for this.  In addition, we apply the same size filter (<700AA's) to the background that we applied to the target in 01_collate_databases.Rmd

```{r}
tg <- read_faa("cache/positive032020_98.fasta") %>%
  add_column(Label = "Tg")

bg <- read_faa("raw_data/swissprot_all_MAY98.fasta") %>%
  add_column(Label = "Bg")

bg <- bg %>% 
  remove_nonstandard_aa() %>%
  filter(between(nchar(seq_aa),20,700)) %>%
  filter(!seq_aa %in% tg$seq_aa) %>%
  sample_n(nrow(tg))

df_to_faa(bg, "cache/negative032020_98.fasta")
```

## Running the tuning script

This needs to be run on an HPC system with R installed and with packages `caret`, `ampir`, `doParallel`, and `R.utils` installed

```bash
module load R/3.6.1
Rscript tune_model.R outfile=tuned.rds target=cache/positive032020_98.fasta background=cache/negative032020_98.fasta ncores=16
```


Or to run on the PBSPro batch system. Edit the tune_model.sh script and then submit using
```bash
qsub tune_model.sh
```

## Read in tuned model

```{r}
tuned_model <- readRDS("cache/tuned.rds")
```

The final values used for the model were sigma = 0.1 and C = 16.

*I think maybe we should modify the tuning values. Mayve C to max 10 ? I worry the large value of C (16) overfits data*
*some links about svm rbf tuning:*

 - https://rstudio-pubs-static.s3.amazonaws.com/310301_37d790cb0eb9455583a0c5d8df6fd472.html
 - https://www.rpubs.com/CHENW05/520528
 - https://towardsdatascience.com/svm-hyper-parameter-tuning-using-gridsearchcv-49c0bc55ce29
 - https://scikit-learn.org/stable/auto_examples/svm/plot_rbf_parameters.html

## Test model

*Need to run script again with either:*

  *1. a line to save the training/testing set*
  
  *2. a line to read in the training/testing set from `create_model.Rmd`*
  
*as it cannot be tested now (as the training/test set for `tuned.rds` is different to the one used for the current `02_create_model.Rmd` training/test set)*
