---
title: "Model tuning"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, eval = TRUE)
```

```{r, echo = FALSE}
library(caret)
```

## Model Tuning

The default model included with ampir is an svm with radial kernel and has two tuning parameters, sigma and C (cost).  Optimal values for these tuning parameters were obtained by using model tuning functions available in `caret`.  Since this is computationally intensive it was performed on an HPC system. The R script `scripts/tune_model.R` provides a convenient wrapper for the required functions and allows it to be run easily on an HPC system. 

This needs to be run on an HPC system with R installed and with packages `caret`, `ampir`, `doParallel`, and `R.utils` installed

```bash
module load R/3.6.1
Rscript tune_model.R outfile=../cache/tuned_precursor.rds train=../cache/featuresTrain_precursor.rds test=../cache/featuresTest_precursor.rds ncores=24
```

Inspecting the results reveals the variation of model performance with the tuning parameters.  If sigma is small and/or cost is small then the model will be more 'local'

```{r}
tuning_data <- readRDS("cache/tuned_precursor.rds")

trellis.par.set(caretTheme())
plot(tuning_data)  

tuning_data$bestTune
```



```{r}
tuning_data$coefnames
```


To keep the final model size small for `ampir` distribution the training is performed with optimal parameters one more time.  This results in a train object with all information required by ampir but without the suboptimal models.

```{r}
#predictors = colnames(tuning_data$trainingData)[-1]

featuresTrain <- readRDS("cache/featuresTrain_precursor.rds")

#predictor_indices <- which(colnames(featuresTrain) %in%  predictors)
predictor_indices <- c(2:28)

tuned_compact_model <- train(Label~.,
      data = featuresTrain[,c(predictor_indices,46)],
      method="svmRadial",
      trControl = trainControl(method = "repeatedcv", number = 10, repeats = 5, classProbs = TRUE),
      preProcess = c("center", "scale"),
      tuneGrid = tuning_data$bestTune)

saveRDS(tuned_compact_model,"cache/ampir_default_model.rds")
```


