---
title: "Model tuning"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, eval = TRUE)
```

```{r, echo = FALSE}
library(caret)
```

The default models included with `ampir` are both SVMs with radial kernel and have two tuning parameters, sigma and C (cost). Optimal values for these tuning parameters were obtained by using model tuning functions available in `caret`. Since this is computationally intensive it was performed on an HPC system. The R script `scripts/tune_model.R` provides a convenient wrapper for the required functions and allows it to be run easily on an HPC system. See the `scripts` directory for details of how this was run.

Inspecting the results for the mature peptide model reveals the variation of model performance with the tuning parameters.  

```{r}
tuning_data <- readRDS("cache/tuned_mature.rds")

trellis.par.set(caretTheme())
plot(tuning_data)
```

Best tune:
```{r}
tuning_data$bestTune
```

And for the precursor model:

```{r}
tuning_data <- readRDS("cache/tuned_precursor_imbal.rds")

trellis.par.set(caretTheme())
plot(tuning_data)
```

Best tune:
```{r}
tuning_data$bestTune
```


