---
title: "Precision-recall curve"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r, echo=FALSE}
library(tidyverse)
library(patchwork)
library(viridis)
source("R/calc_cm_metrics.R")
```

Read in predicted probabilities data
```{r}
ampir_prob_data <- readRDS("ampir_0.1.0_data/ampir_prob_data.rds")

```

Calculate performance metrics over a range of probabilities 
```{r}
ampir_roc_data <- as.data.frame(t(sapply(seq(0.01, 0.99, 0.01), calc_cm_metrics, ampir_prob_data)))
```


### Theoretical AMP content

A new variable, $\alpha$, was introduced to represent the percentage of AMPs in the test set to more easily create precision-recall curves for various AMP proportions. The calculation for the recall metric for $\alpha$ remains the same but the precision metric has been slightly modified (see below):


$$Precision_{\alpha} = \frac{TP\alpha}{TP\alpha + FP(1-\alpha)}$$

$$Recall_{\alpha} = \frac{TP}{TP + FN}$$

Function that uses the recall-precision metric calculations for any $alpha$ value
```{r}
calc_precision_recall <- function(df,alpha) {
  df %>% 
  mutate(Recall = Recall) %>% 
  mutate(Precision = TP*alpha / (TP*alpha+FP*(1-alpha))) %>% 
  select(Recall,Precision,p_threshold)
}
```

Use the function for a range of $alpha$ values and collapse to data frame

```{r}
pr_data <- do.call(rbind,lapply(c(0.01,0.05,0.1,0.5),function(alpha) {
  calc_precision_recall(ampir_roc_data,alpha) %>% add_column(alpha=alpha)
}))
```

Plot using a traditional precision vs recall curve.  This is useful in the sense that it very clearly shows the tradeoff between the two. A useful way to think of Precision is that it defines the "Purity" of our predicted set of AMPs whereas the Sensitivity or Recall defines the "Completeness" of the predicted AMP set.  We want to choose the p_threshold so that there is a balance or Purity and Completeness.  When `alpha` is high this is easy to do, but when it is low it becomes a very difficult tradeoff.

```{r}
ggplot(pr_data, aes(x=Recall, y=Precision)) +
  geom_line(aes(group=as.factor(alpha), colour = as.factor(alpha))) + 
  scale_color_viridis(discrete = TRUE) +
  theme(legend.key = element_rect(fill = "white"),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "grey")) +
  labs(colour = "alpha") +
   guides(color = guide_legend(reverse=TRUE))
```


```{r, echo=FALSE, eval=FALSE}
ggsave("figures/alpha_gradient.png", height = 5, width = 7)
```


Plot an explicit axis for `p_threshold`. This is useful for choosing the threshold value. Also note that as $\alpha$ gets smaller and smaller the Precision curve shifts so that high values of precision are only achieved for very high `p_threshold` values.

```{r}
pr_data_long <- pr_data %>% gather("metric","value",-p_threshold,-alpha)


variable_names <- c("0.01" = "Proportion of AMPs in genome: 0.01",
                    "0.05" = "Proportion of AMPs in genome: 0.05",
                    "0.1" = "Proportion of AMPs in genome: 0.10",
                    "0.5" = "Proportion of AMPs in genome: 0.50")
```


```{r}
ggplot(pr_data_long,aes(x=p_threshold,y=value)) + 
  geom_line(aes(linetype=metric)) + facet_wrap(~alpha, labeller= as_labeller(variable_names)) +
  labs(x = "Probability threshold", y = "", linetype = "") +
  scale_x_continuous(breaks=c(0, 0.50, 1.00)) +
  scale_y_continuous(breaks=c(0, 0.50, 1.00)) +
  theme(legend.position = "left",
        legend.text = element_text(angle = 90, hjust = 0.5),
        legend.margin = margin(0,0,0,8),
        legend.box.margin=margin(-10,-10,-10,-10),
        legend.key = element_rect(fill = "white"),
        legend.key.height = unit(2, "cm"),
        legend.box.spacing = unit(0.1, "cm"),
        panel.background = element_blank(),
        axis.line = element_line(colour = "grey")) +
  guides(linetype = guide_legend(reverse=TRUE))
```


```{r, eval = FALSE, echo=FALSE}
ggsave("figures/alphas.png", width = 7, height=5)
```

Extract 1% alpha value
```{r}
pr_data_alpha1_long <- pr_data_long %>% filter(alpha == 0.01)
```

```{r, eval = FALSE, echo=FALSE}
ggplot(filter(pr_data_alpha1_long, p_threshold >= 0.5), aes(x=p_threshold, y=value)) + 
  geom_line(aes(colour = metric)) + 
  labs(x = "Probability threshold", y = "", colour = "") +
  scale_colour_manual(values = c("blueviolet", "forestgreen")) +
  scale_x_continuous(breaks=c(0.50, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  scale_y_continuous(breaks=c(0.10, 0.50, 1.00)) +
  theme(legend.position = c(0.2, 0.5),
        panel.background = element_blank(),
        legend.key=element_blank(),
        axis.line = element_line(colour = "grey")) +
  guides(colour = guide_legend(reverse = TRUE))

```



```{r}
alpha1 <- ggplot(filter(pr_data_alpha1_long, p_threshold >= 0.5), aes(x=p_threshold, y=value)) + 
  geom_line(aes(colour = metric)) + 
  labs(x = "Probability threshold", y = "", colour = "") +
  scale_colour_manual(values = c("blueviolet", "forestgreen")) +
  scale_x_continuous(breaks=c(0.50, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  theme(legend.position = c(0.2, 0.5),
        panel.background = element_blank(),
        legend.key=element_blank(),
        axis.line = element_line(colour = "grey")) +
  guides(colour = guide_legend(reverse = TRUE)) +
  ggtitle("B")
```

### From `benchmark_auroc.Rmd`

Read performance metrics for different models over a range of range of probability thresholds
```{r}
models_roc <- readRDS("cache/models_roc.rds")
```

```{r}
benchmark_roc <- ggplot(models_roc) + 
  geom_line(aes(x = FPR, y = Recall, colour = model)) + 
  xlim(0,1) +
  labs(x = "False positive rate", y = "True positive rate", colour = "Model and AUC") +
  scale_colour_manual(breaks= c("ampir", "amp_scanner", "amPEP", "iAMPpred"),
                      labels=c("ampir - 96%", "amp scanner - 86%", "amPEP - 67%", "iAMPpred - 58%"),
                      values = c("blueviolet", "goldenrod2", "blue4", "cyan")) +
  theme(legend.position = c(0.76, 0.25),
        legend.key=element_blank(),
        legend.title = element_text(size = 8),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "grey")) +
  ggtitle("A")
```


Combine the AUROC benchmark and the 1% alpha plots 

```{r}
benchmark_roc | alpha1
roc_and_alpha1 <- benchmark_roc | alpha1
```


```{r, eval = FALSE, echo=FALSE}
ggsave("figures/roc_and_alpha1.jpg", roc_and_alpha1, width = 7, height=4)

```

